*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-01 Code/Manuscript_O2.Rmd
Added Amanda's MIT9313 explanation for O2 and respiration

Under full atmospheric [O~2~] and blue light, LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed [@aroPhotoinhibitionPhotosystemII1993], producing damaging singlet oxygen (^1^O~2~) [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017; @murphyPhotoinactivationPhotosystemII2017; @hakalaEvidenceRoleOxygenevolving2005]. The presence of NATL2A, a clade LLI, has been found near the ocean surface during deep ocean mixing [@malmstromTemporalDynamicsProchlorococcus2010]. Malmstrom *et al*. attributes NATL2a tolerance to short exposures of high light to the presence of the genes encoding photolyase, a gene found in HL clades which, in the presence of blue light, is responsible for repairing DNA damaged by UV light [@sancarStructureFunctionDNA2003]. Repair of photoinactivated PSII relies on the removal of damaged PsbA [@mannInvolvementFtsHHomologue2000; @adamFtsHProteasesChloroplasts2005], followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. Degradation of PsbA is a rate-limiting step in recovery from photoinhibition [@kanervoD1ProteinDegradation1993], mediated largely by a heterohexamer of (FtsH12)~3~, a membrane-bound (Sacharz et al., 2015)(Zak et al., 2001) metalloprotease  (Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014).  [@adamFtsHProteasesChloroplasts2005; @boehmSubunitOrganizationSynechocystis2012a; @nixonRecentAdvancesUnderstanding2010; @komendaExposedNTerminalTail2007] (Pisareva et al., 2007; Sacharz et al., 2015),  @mannInvolvementFtsHHomologue2000, (AMANDA REF from ALGATECH).

@@

West *et al*. [@westNichePartitioningProchlorococcusPopulationsStratified1999] and Malmstrom *et al*. [@malmstromTemporalDynamicsProchlorococcus2010] found that decreased abundances of the LL clades corresponded to increased depth of the surface mixed layer. Malmstrom *et al*. [@malmstromTemporalDynamicsProchlorococcus2010] attributes the transport of LL ecotypes to the surface and consequent exposure to photoinhibitory high light levels as the reason for low cell abundances with increased mixed layer depth. @westNichePartitioningProchlorococcusPopulationsStratified1999 found the depth of the mixed layer strongly influenced the depth transition from HL to LL clades, but that factors other than light levels may influence the variations in the upper and lower depth limits of these ecotypes. We hypothesize that low cell abundances of LL ecotypes in the mixed layer is likely driven in part by increased [O~2~],  and it is [O~2~] that constrains LL clades to deeper waters, not necessarily the light level. We found that under 25 µM O~2~ representatives of 'LL' clades, SS120 and MIT9313, actually tolerate approximately 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of PUR [@morelAvailableUsableStored1978],  comparable to the representative HL clade, MED4 which also exhibited growth saturation at the same cumulative diel PUR of 1.0 x 10^6^ µmol photons m^-2^ d^-1^.  Growth under lower O~2~ allowed MIT9313 to substantially increase its exploitation of higher diel PUR.

@@

>>>>

Under full atmospheric [O~2~] and blue light, LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed [@aroPhotoinhibitionPhotosystemII1993], producing damaging singlet oxygen (^1^O~2~) [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017; @murphyPhotoinactivationPhotosystemII2017; @hakalaEvidenceRoleOxygenevolving2005]. The presence of NATL2A, a clade LLI, has been found near the ocean surface during deep ocean mixing [@malmstromTemporalDynamicsProchlorococcus2010]. Malmstrom *et al*. attributes NATL2a tolerance to short exposures of high light to the presence of the genes encoding photolyase, a gene found in HL clades which, in the presence of blue light, is responsible for repairing DNA damaged by UV light [@sancarStructureFunctionDNA2003]. Repair of photoinactivated PSII relies on the removal of damaged PsbA [@mannInvolvementFtsHHomologue2000; @adamFtsHProteasesChloroplasts2005], followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. Degradation of PsbA is a rate-limiting step in recovery from photoinhibition [@kanervoD1ProteinDegradation1993], mediated largely by a heterohexamer of (FtsH12)~3~, a membrane-bound [@sacharzSubCellularLocation2015; @zakInitialStepsBiogenesis2001] metalloprotease  [(Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014).  [@adamFtsHProteasesChloroplasts2005; @boehmSubunitOrganizationSynechocystis2012a; @nixonRecentAdvancesUnderstanding2010; @komendaExposedNTerminalTail2007] (Pisareva et al., 2007; Sacharz et al., 2015),  @mannInvolvementFtsHHomologue2000, (AMANDA REF from ALGATECH).

@@

West *et al*. [@westNichePartitioningProchlorococcusPopulationsStratified1999] and Malmstrom *et al*. [@malmstromTemporalDynamicsProchlorococcus2010] found that decreased abundances of the LL clades corresponded to increased depth of the surface mixed layer. Malmstrom *et al*. [@malmstromTemporalDynamicsProchlorococcus2010] attributes the transport of LL ecotypes to the surface and consequent exposure to photoinhibitory high light levels as the reason for low cell abundances with increased mixed layer depth. West *et al*.  [@westNichePartitioningProchlorococcusPopulationsStratified1999] found the depth of the mixed layer strongly influenced the depth transition from HL to LL clades, but that factors other than light levels may influence the variations in the upper and lower depth limits of these ecotypes. We hypothesize that low cell abundances of LL ecotypes in the mixed layer is likely driven in part by increased [O~2~],  and it is [O~2~] that constrains LL clades to deeper waters, not necessarily the light level. We found that under 25 µM O~2~ representatives of 'LL' clades, SS120 and MIT9313, actually tolerate approximately 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of PUR [@morelAvailableUsableStored1978],  comparable to the representative HL clade, MED4 which also exhibited growth saturation at the same cumulative diel PUR of 1.0 x 10^6^ µmol photons m^-2^ d^-1^.  Growth under lower O~2~ allowed MIT9313 to substantially increase its exploitation of higher diel PUR.

@@

MIT9313 is not “using” photosynthesis in low O~2~ environments to do CO~2~ fixation so much as to generate oxygen locally in an environment that is low enough to limit the ability to do respiration. The deep water environments that we see MIT 9313 are relatively nutrient rich and there is evidence in the literature of *Prochlorococcus* have  been observed to take up and metabolise various sugars [@gomez-baenaGlucoseUptakeIts2008; @munoz-marinGlucoseUptakeProchlorococcus2017] and amino acids (Zubkov et al., 2004). They may be using photosynthesis to generate O~2~ to act as an electron sink for respiration. And maybe using ATP to take up food from the surroundings.
Partensky *et al*. [@partenskyComparisonPhotosyntheticPerformances2018] found that in the low-light conditions found in the OMZ, MED4, SS120 and MIT9313 all became net O~2~ consumers, suggesting that the low light levels caused the respiratory chain to consume more O~2~ than the photosynthetic electron transport chain could generate and could potentially be contributing to the maintenance of the low O~2~ environment
